---
title: "phase_1"
author: "Dinah"
date: "2025-07-03"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

**Loading the data set.**

```{r}
library(readr)
my_dataset <- read_delim("C:/Users/Asus/Downloads/Quiet-quitting (4)/Quiet-quitting/quiet_quitting_dataset.psv", 
    delim = "|", escape_double = FALSE, trim_ws = TRUE)
View(my_dataset)
```

**remove null values**

```{r}
my_dataset <- na.omit(my_dataset)

```

**loading libraries**

```{r}
library(bayesrules)
library(tidybayes)
library(bayesplot)
library(rstanarm)
library(tidyverse)
library(broom.mixed)
library(readr)
library(readxl)


```

**Description of univariate distributions.**

```{r}
summary(my_dataset)
str(my_dataset)

```

```{r}
glimpse(my_dataset)

```

```{r}
colnames(data)
```

```{r}
library(dplyr)

numeric_vars <- my_dataset %>% select(where(is.numeric))
categorical_vars <- my_dataset %>% select(where(is.character))


```

**Histogram for numeric values**

```{r}
library(ggplot2)
library(tidyr)

numeric_vars %>%
  pivot_longer(everything()) %>%
  ggplot(aes(x = value)) +
  geom_histogram(bins = 30, fill = "skyblue", color = "black") +
  facet_wrap(~ name, scales = "free") +
  theme_minimal() +
  labs(title = "Histograms of Numeric Variables")


```

**Description of bivariate distributions.**

```{r}
outcome_var <- "Quiet_quitting_bahaviour"

```

**Job satisfaction vs Quiet quitting.**

```{r}
ggplot(my_dataset, aes(x = Job_satisfaction, y = Quiet_quitting_behavior)) +

  geom_point(alpha = 0.3) +

  geom_smooth(method = "lm", color = "red") +

  labs(title = "Job Satisfaction vs Quiet Quitting", x = "Job Satisfaction", y = "Quiet Quitting Behavior")


```

**Perceived_org_support vs quiet quitting**

```{r}
ggplot(my_dataset, aes(x = Perceived_org_support , y = Quiet_quitting_behavior)) +
  geom_point(alpha = 0.3) +
  geom_smooth(method = "lm", color = "red") +
  labs(title = "Perceived_org_support  vs Quiet Quitting", x = "Perceived_org_support ", y = "Quiet Quitting Behavior")

```

```{r}
ggplot(my_dataset, aes(x = Work_engagement, y = Quiet_quitting_behavior)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "lm", se = FALSE, color = "red") +
  labs(title = "Work Engagement vs Quiet Quitting Behavior",
       x = "Work Engagement", y = "Quiet Quitting Behavior")
```

```{r}
ggplot(my_dataset, aes(x = Psychological_meaningfulness, y = Quiet_quitting_behavior)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "lm", se = FALSE, color = "red") +
  labs(title = "Psychological Meaningfulness vs Quiet Quitting Behavior",
       x = "Psychological Meaningfulness", y = "Quiet Quitting Behavior") +
  theme_minimal()

```

```{r}
ggplot(my_dataset, aes(x = Org_culture, y = Quiet_quitting_behavior)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "lm", se = FALSE, color = "orange") +
  labs(title = "Organizational Culture vs Quiet Quitting Behavior",
       x = "Org Culture", y = "Quiet Quitting Behavior") +
  theme_minimal()

```

```{r}
ggplot(my_dataset, aes(x = Psychological_resilience, y = Quiet_quitting_behavior)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "lm", se = FALSE, color = "brown") +
  labs(title = "Psychological Resilience vs Quiet Quitting Behavior",
       x = "Psychological Resilience", y = "Quiet Quitting Behavior") +
  theme_minimal()


```

```{r}
ggplot(my_dataset, aes(x = moderation_term, y = Quiet_quitting_behavior)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "lm", se = FALSE, color = "red") +
  labs(title = "Moderation Term vs Quiet Quitting Behavior",
       x = "Moderation Term", y = "Quiet Quitting Behavior") +
  theme_minimal()

```

**Identify any anomalous and outlier observations along with associated rationale for identifying them as such**

```{r}

numeric_data <- my_dataset %>% select(where(is.numeric))  
iqr_outliers <- function(x) {
  Q1 <- quantile(x, 0.25)
  Q3 <- quantile(x, 0.75)
  IQR <- Q3 - Q1
  lower <- Q1 - 1.5 * IQR
  upper <- Q3 + 1.5 * IQR
  return(which(x < lower | x > upper))
}

iqr_results <- lapply(numeric_data, iqr_outliers)

# Print count of outliers per variable
sapply(iqr_results, length)


```

```{r}
z_score_outliers <- function(x) {
  z <- scale(x)
  which(abs(z) > 3)
}

z_results <- lapply(numeric_data, z_score_outliers)

# Print count of outliers per variable
sapply(z_results, length)

```

```{r}
numeric_data %>%
  pivot_longer(cols = everything()) %>%
  ggplot(aes(x = name, y = value)) +
  geom_boxplot(fill = "lightblue", outlier.color = "red") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(title = "Boxplot of Numeric Variables (Outliers Highlighted)",
       x = "Variables", y = "Values")



```

**Specify a comprehensive model that includes all of the predictors and the outcome.**

```{r}
full_model <- stan_glm(
  Quiet_quitting_behavior ~ .,
  data = my_dataset,
  family = gaussian(),         # continuous outcome
  prior = normal(0, 2.5),      # weakly informative prior
  prior_intercept = normal(30, 5),  # prior around the center of your outcome
  chains = 4,
  iter = 2000,
  seed = 123
)

```

```{r}
summary(full_model)        
     

```

```{r}
mcmc_trace(full_model)
```

```{r}
mcmc_dens_overlay(full_model)
```

```{r}
mcmc_acf(full_model)
```

```{r}
neff_ratio(full_model)
```

```{r}
# Get credible intervals & hypothesis testing
posterior_interval(full_model, prob = 0.95)

```

```{r}
rhat(full_model)
```

```{r}
tidy(full_model,
     effects=c("fixed","aux"),
     conf.int = TRUE,
     conf.level = 0.95) ->output
as.data.frame(output)
     
```
